Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-26 VMCAI-revised-version/final.tex
up

This all changes when one considers VASS (having control-states). For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with control-state $q$ \----
then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none. If there is none, then 
$\mathop{\uparrow} q(\textbf{0})$
 is in the complement of $\Bad$, i.e. $\Safe$.
 From then on, we partition $\Safe$ into two subsets: an upward-closed 
union of sets of the form $\mathop{\uparrow} q(\textbf{0})$, and a remaining downward-closed subset.
The key now is that
there is only a finite number of elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} \leq \textbf{u}_k$ for all of which we will check reachability of $\Safe$ in $k$ or less steps, while 
 elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps. 
Only being concerned by the upward-closed subset of $\Safe$ in the latter case enable us to reuse techniques seen when $\Safe$ is upward-closed, in particular we can compute the basis for the sets of elements from which the upward-closed subset of $\Safe$ is reachable in at most $k$ steps. 
Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$. This lead to a decision procedure detailed in appendix~\ref{appendix vass},
 which lead to the following decidability result:
\begin{theorem}\label{vass down}
{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
$\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}

>>>>

This all changes when one considers VASS (having control-states). For a VASS $V = (Q,T)$, a set $\Bad = \mathop{\uparrow} \Bad$, for all $q \in Q$, either there is an element of the basis of $\Bad$ with control-state $q$ \----
then $q(\textbf{v})$ with $\textbf{v} > \textbf{v}_{\Bad}$ is necessarily in $\Bad$ by upward closure; if this hold for all $q \in Q$, then $k$-resilience does not hold \---- either there is none. If there is none, then 
$\mathop{\uparrow} q(\textbf{0})$
 is in the complement of $\Bad$, i.e. $\Safe$.
Based on this, we partition $\Safe$ into two subsets: an upward-closed 
union of sets of the form $\mathop{\uparrow} q(\textbf{0})$, and a remaining downward-closed subset.
The key now is that
there is only a finite number of elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} \leq \textbf{u}_k$ for all of which we will check reachability of $\Safe$ in $k$ or less steps, while 
 elements of $\Bad$ of the form $p(\textbf{v})$ with $ \textbf{v} > \textbf{u}_k$ can only potentially reach the upward-closed subset of $\Safe$ in $k$ or less steps. 
Only being concerned by the upward-closed subset of $\Safe$ in the latter case enable us to reuse techniques seen when $\Safe$ is upward-closed. Indeed we can compute the basis for the sets of elements from which our upward-closed subset of $\Safe$ is reachable in at most $k$ steps. 
Substracting these predecessor from $\Bad$ yields either a finite number of elements from which one has to check $\Safe$ is reachable in at most $k$ steps, either an infinite number of elements of which there is one which cannot reach $\Safe$ in at most $k$ steps \-- for much the same reasons {\sc Bounded resilience} and {\sc $k$-resilience} never hold for VAS when $\Safe = \mathop{\downarrow} \Safe$ and $\Bad = \mathop{\uparrow} \Bad$. This leads to a decision procedure detailed in appendix~\ref{appendix vass},
 which lead to the following decidability result:
\begin{theorem}\label{vass down}
{\sc $k$-resilience }  and {\sc Bounded resilience} are decidable for VASS when 
$\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}
